---
title: "STAT 231: Problem Set 7B"
author: "Alastair Poole"
date: "due by 10 PM on Friday, April 30"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

This homework assignment is designed to help you further ingest, practice, and expand upon the material covered in class over the past week(s).  You are encouraged to work with other students, but all code and text must be written by you, and you must indicate below who you discussed the assignment with (if anyone).  

Steps to proceed:

\begin{enumerate}
\item In RStudio, go to File > Open Project, navigate to the folder with the course-content repo, select the course-content project (course-content.Rproj), and click "Open" 
\item Pull the course-content repo (e.g. using the blue-ish down arrow in the Git tab in upper right window)
\item Copy ps7B.Rmd from the course repo to your repo (see page 6 of the GitHub Classroom Guide for Stat231 if needed)
\item Close the course-content repo project in RStudio
\item Open YOUR repo project in RStudio
\item In the ps7B.Rmd file in YOUR repo, replace "YOUR NAME HERE" with your name
\item Add in your responses, committing and pushing to YOUR repo in appropriate places along the way
\item Run "Knit PDF" 
\item Upload the pdf to Gradescope.  Don't forget to select which of your pages are associated with each problem.  \textit{You will not get credit for work on unassigned pages (e.g., if you only selected the first page but your solution spans two pages, you would lose points for any part on the second page that the grader can't see).} 
\end{enumerate}

```{r, setup, include=FALSE}
library(tidyverse)
library(viridis)
library(readr)
# for networks
library(ggnetwork)
library(igraph)
# for mapping
library(datasets)
library(mdsr)
library(gapminder)
library(fivethirtyeight)

knitr::opts_chunk$set(
  tidy=FALSE,     # display code as typed
  size="small")   # slightly smaller font for code
```


\newpage 
# If you discussed this assignment with any of your peers, please list who here:

> ANSWER: Ethan Gilman, Brandon Kwon

\newpage
# 1. More migration

1a. Consider migration between the following countries: Argentina, Brazil, Japan, Kenya, Great Britain, India, South Korea, United States.  Compare the TOTAL (males + females) migration between these countries over time.  In separate (directed) graphs for 1980 and 2000, visualize the network for the these countries with edge width and/or edge color corresponding to migration flow size.  Interpret the two graphs -- what *information in context* do they convey?

*Don't forget to order the columns correctly and only keep relevant rows before transforming into a network object.*

> ANSWER: Generally, it appears as though there was much more total migration in the year 1980. This is seen through the high quantity of thicker edges, and the edges themselves appear more abundant that in the year 2000. In 1980, the USA and Great Britain appeared to be the two most central vertices, which would mean that the most number of people were migrating either to or from those countries. This makes sense in context because during the 1980s, many people viewed the prosperity in the US and the UK as opportune, and so they moved there. However, in 2000 Japan and Korea also became much more involved in migration flows. This makes sense due to the rapid industrialization of these two countrie resulting in many more people worldwide viewing these countries as viable destination locations.


```{r, message = FALSE}
path_in <- "~/Desktop/Data-Science-231/Labs"
MigrationFlows <- read_csv(paste0(path_in,"/MigrationFlows.csv"))

# Argentina, Brazil, Great Britain, Japan, Kenya,  India, South Korea, United States
countries <- c("ARG", "BRA", "GBR", "JPN", "KEN", "IND","KOR", "USA")

# need migration overall: 
# do some prelim data wrangling to combine numbers for males + females

#Total migration for 1980
MigrationFlows2 <- MigrationFlows %>%
  filter(destcode %in% countries & origincode %in% countries, Y1980 > 0) %>%
  group_by(destcode, origincode) %>%
  summarize(total1980 = sum(Y1980))

#Total migration for 2000
MigrationFlows3 <- MigrationFlows %>%
  filter(destcode %in% countries & origincode %in% countries, Y2000 > 0) %>%
  group_by(destcode, origincode) %>%
  summarize(total2000 = sum(Y2000))

migration1980 <- graph_from_data_frame(MigrationFlows2, directed = TRUE)
migration2000 <- graph_from_data_frame(MigrationFlows3, directed = TRUE)

migration_network1980 <- ggnetwork(migration1980)

ggplot(data = migration_network1980
       , aes(x = x, y = y, xend = xend, yend = yend)) +
  geom_edges(arrow=arrow(type="closed", length=unit(6,"pt"))
            , aes(size = total1980)) +
  geom_nodes() +
  geom_nodelabel(aes(label = name)) +
  theme_blank()

migration_network2000 <- ggnetwork(migration2000)

ggplot(data = migration_network2000
       , aes(x = x, y = y, xend = xend, yend = yend)) +
  geom_edges(arrow=arrow(type="closed", length=unit(6,"pt"))
            , aes(size = total2000)) +
  geom_nodes() +
  geom_nodelabel(aes(label = name)) +
  theme_blank()
```

1b. Compute the *unweighted* in-degree for Japan in this network from 2000, and the *weighted* in-degree for Japan in this network from 2000.  In 1-2 sentences, interpret these numbers in context (i.e., without using the terms "in-degree" or "weighted").

> ANSWER: These numbers mean that people from the other 7 countries observed migrated into Japan, and a total of 931,809 people migrated into Japan in the year 2000.

```{r}
#unweighted in-degree for Japan
igraph::degree(migration2000, mode = "in")

#weighted in-degree for Japan
strength(migration2000, weights = E(migration2000)$total2000)
```


1c. Among these same countries, identify the top 5 countries *of origin* and *of destination* (separately) in 1980 using (weighted) degree centrality.  Interpret this information.

> ANSWER: The top five countries of origin are USA, Japan, Great Britain, Brazil, and Argentina. This means that these are countries with the most number of people traveling out of in 1980. The top five countries of destination are Korea, Great Britain, India, Japan, and USA This means that these are countries with the most number of people traveling into it in 1980.

```{r}
sort(strength(migration1980, weights=E(migration1980)$total1980, mode="in"))

sort(strength(migration1980, weights=E(migration1980)$total1980, mode="out"))
```


1d. Among these same countries, idenitify the top 5 countries *of origin* and *of destination* (separately) in 2000 using (weighted) degree centrality.  Interpret this information.

> ANSWER: The top five countries of origin are USA, Japan, Great Britain, Brazil, and Argentina. This means that these are countries with the most number of people traveling out of in 2000. The top five countries of destination are Great Britain, Korea, Japan, India, and USA This means that these are countries with the most number of people traveling into it in 2000.

```{r}
sort(strength(migration2000, weights=E(migration2000)$total2000, mode="in"))

sort(strength(migration2000, weights=E(migration2000)$total2000, mode="out"))
```


1e. What is the diameter of this network in 2000? In 1-2 sentences, interpret this value.

> ANSWER: The diameter of this network in 2000 is 2, meaning that in the year 2000, the two countries that were the least connnected in terms of migration flow, were only separated by one other country.

```{r}
diameter(migration2000)
```


1f. What is the density of this network in 2000? In 1-2 sentences, interpret this value.

> ANSWER: The density of this network in 2000 is roughly 0.82. This means that out of all the possible connections possible in this network between the countries in question, 82% of them exist in the network, and 18% of possible connections are not occurring.

```{r}
graph.density(migration2000)
```


\newpage
# 2. Mapping spatial data 

Reproduce the map you created for Lab08-spatial (and finish it if you didn't in class). In 2-4 sentences, interpret the visualization.  What stands out as the central message?  

NOTE: you do NOT need to say what colors are representing what feature (e.g, NOT:  "In this map, I've colored the countries by GDP, with green representing low values and red representing high values") -- this is obvious to the viewer, assuming there's an appropriate legend and title.  Rather, what *information* do you extract from the visualization? (e.g., "From the choropleth below, we can see that the percent change in GDP per capita  between 1957-2007 varies greatly across countries in Central America.  In particular, Panama and Costa Rica stand out as having GDPs per capita that increased by over 200% across those 50 years.  In contrast, Nicaragua's GDP per capita decreased by a small percentage during that same time span.")

> ANSWER: Interestingly, it appears as though the Southeast Region of the US generally has the least number of hate crimes per 100,000. Generally speaking, it also appears as though there are more hate crimes where there are more people living, and this makes sense because if there are more people, then it is likely that there will be more crimes in general. We can also see that Massachusetts, New Jersey, Kentucky, and North Dakota had the most number of hate crimes per 100,000, while in contrast, Wyoming, Georgia, and Pennsylvania seem to have had the least number of hate crimes per 100,000.

```{r}
# states in the US
usa_states <- map_data(map = "state", region = ".") 
#change all states to lowercase
hate_crimes <- hate_crimes %>%
  mutate("state" = tolower(state))
#creating new datset
hate_crimes_map <- hate_crimes %>%
  inner_join(usa_states, by = c("state" = "region"))

ggplot(hate_crimes_map, aes(x = long, y = lat, group = group
                      , fill = avg_hatecrimes_per_100k_fbi)) +
  geom_polygon(color = "white") +
  theme_void() +
  coord_fixed(ratio = 1.3) +
  labs(title = "Average Hate Crimes per 100k, by state*"
       , subtitle = "as Reported by the FBI"
       , caption = "* Hawaii and Alaska not shown above"
       , fill = "") +
  scale_fill_viridis(option = "plasma", direction = -1)
```

\newpage
# 3. Mapping spatial data at a different level

Create a map at the world, country, or county level based on the choices provided in lab08-spatial, that is at a DIFFERENT level than the map you created for the lab (and included above).  For instance, if you created a map of US counties for the lab, then choose a country or world map to create here.

Note: While I recommend using one of the datasets provided in the lab so you don't spend a lot of time searching for data, you are not strictly required to use one of those datasets.  

Describe one challenge you encountered (if any) while creating this map.

> ANSWER: One challenge was how to join the 2 datasets, specifically the initial data wrangling required to allow "inner-join" to be used. Once I completed this, it also took me awhile to figure out that I needed to turn the "area_name" all lowercase, so that it would match the observations in the usa_counties dataset.

```{r}
county_employment <- readxl::read_excel(paste0("~/Desktop/course-content/data/Unemployment.xls")
                                        , sheet = 1
                                        , skip = 7) %>%
  janitor::clean_names()

usa_counties <- map_data(map = "county", region = ".")

county_employment <- county_employment %>% 
  mutate(area_name = gsub(pattern = " County.*", replacement = "", x = area_name),
         area_name = gsub(pattern = " Borough.*", replacement = "", x = area_name),
         area_name = gsub(pattern = " Area.*", replacement = "", x = area_name),
         area_name = gsub(pattern = " Municipality.*", replacement = "", x = area_name),
         area_name = str_to_lower(area_name))

final_df <- usa_counties %>%
  inner_join(county_employment, by = c("subregion" = "area_name"))

 
ggplot(final_df, aes(x = long, y = lat, group = group, fill = unemployment_rate_2000)) +
  geom_polygon(colour = "white") +
  theme_void() +
  coord_fixed(ratio = 1.3)+
  labs(title = "Unemployment Rate by Counties in the United States*"
       , subtitle = "In the Year 2000"
       , caption = "* Hawaii and Alaska not shown above"
       , fill = "") +
  scale_fill_viridis(option = "magma", direction = -1)

```


\newpage
# 4. Love Actually (OPTIONAL PRACTICE)

This problem is *optional* and will not be graded, but is given to provide additional practice interpreting networks and as another real-world example of network analysis that might be intriguing to film buffs. 

Consider the figure "The Two Londons of 'Love Actually'" in [this FiveThirtyEight article](https://fivethirtyeight.com/features/the-definitive-analysis-of-love-actually-the-greatest-christmas-movie-of-our-time/). 

2a. Based on this figure, is the network connected?  In 1-2 sentences, please explain.

> ANSWER: 

2b. Based on the figure, what is the (unweighted) degree for Emma Thompson?  What is the (unweighted) degree for Keira Knightley?  Explain what these values mean for these characters.

> ANSWER: 

2c. Based on the figure, for whom would the (unweighted) betweenness centrality measure be higher: Colin Firth or Hugh Grant?  Explain what this implies. 

> ANSWER:  


\newpage
# 5. Migration network on a world map! (OPTIONAL PRACTICE) 

This problem is *optional* and will not be graded, but is given to provide additional coding practice and as a challenge to incorporate networks and mapping techniques together.   

Create a world map that visualizes the network of countries we examined in #1 for the year 2000.  For example, arrows to and from each of countries on the world map could have edge widths relative to their weighted degree centrality to represent migration to and from the countries.

Code to get you started is provided below.

```{r}
# from mdsr package
# should see 'world_cities` df in your environment after running
data(world_cities)

# two-letter country codes
# Argentina, Brazil, Great Britain, Japan, Kenya
# India, South Korea, United States
countries2 <- data.frame(country3=countries
                        , country2 = c("AR", "BR", "GB", "JP"
                                         , "KE", "IN","KR", "US"))

# find capitals for anchoring points; can't find D.C., use Boston
cities <- c("Buenos Aires", "Brasilia", "London", "Tokyo", "Nairobi"
             , "New Delhi", "Seoul", "Boston")

anchors <- world_cities %>%
  right_join(countries2, by = c("country" = "country2")) %>%
  filter(name %in% cities) %>%
  select(name, country, country3, latitude, longitude) 

# one suggested path:
# 1. based on the anchors dataset above and your Migration 2000 dataset created for # 1, 
#    create dataframe that would supply geom_curve with the relevant arrow locations
#    (start points and end points)
# 2. create world map dataset using `map_data` function
# 3. use geom_polygon to create world map, geom_point and/or geom_text to add 
#    city points, and geom_curve to add weighted/colored arrows


```

